---
title: "pkgdiff"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pkgdiff}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pkgdiff)
```
The purpose of **pkgdiff** is to identify changes between package versions,
with the aim of assessing package stability.  Here is a list of the major
functions in **pkgdiff**, and a short explanation of each:

- [**pkg_info()**](#pkg_info): Displays general information about a package,
including a list of all exported functions.
- [**pkg_diff()**](#pkg_diff): Compares two versions of a package, and shows
which functions have been added or removed.
- [**pkg_stability()**](#pkg_stability): Looks at the entire history of a package,
identifies breaking changes, and assesses the package for stability.
- [**pkg_repo()**](#pkg_repo): Generates a list of the packages in a repository
and their versions.
- [**repo_breakages()**](#repo_breakages): Looks at a package repository and tries to identify
which packages will break when upgraded.
- [**repo_stability()**](#repo_stability): Looks at a package repository and calculates
a stability score for each package.

Below is a explanation and demo for each of the above functions.

## General Package Information: pkg_info(){#pkg_info}


## Comparing Package Versions: pkg_diff(){#pkg_diff}


## Stability Assessment: pkg_stability(){#pkg_stability}


## Listing Package Versions: pkg_repo(){#pkg_repo}

## Anticipating Breakages: repo_breakages(){#repo_breakages}

## Assessing Entire Repository: repo_stability(){#repo_stability}


