---
title: "pkgdiff"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pkgdiff}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pkgdiff)
```
The purpose of **pkgdiff** is to identify changes between package versions,
with the aim of assessing package stability.  Here is a list of the major
functions in **pkgdiff**, and a short explanation of each:

- `pkg_info()`: Displays general information about a package,
including a list of all exported functions.
- `pkg_stability()`: Looks at the history of a package,
identifies breaking changes, and assesses the package for stability.
- `pkg_diff()`: Compares two versions of a package, and shows
which functions have been added or removed.
- `pkg_repo()`: Generates a list of the packages in a repository
and their versions.
- `repo_breakages()`: Looks at a package repository and tries to identify
which packages will break when upgraded.
- `repo_stability()`: Looks at a package repository and calculates
a stability score for each package.

Below is a short explanation and demo for each of the above functions.

## General Package Information

The firt thing you may want to do when exploring a package is to get
some general information about it.  This general package information
can be achieved using the `pkg_info()` function.  All you need to 
pass is the package name, and the function will return basic 
information such as the maintainer, description,
and imported packages.  Here is a simple demonstration:

```{r eval=FALSE, echo=TRUE}
library(pkgdiff)

# Examine fmtr package
pkg_info("fmtr")
# # A package info object: fmtr package
# - Version: v1.6.5
# - Release Date: 2024-06-13
# - Title: Easily Apply Formats to Data
# - Maintainer: David Bosak <dbosak01@gmail.com>
# - License: CC0
# - Description: Contains a set of functions that can be used to apply
# formats to data frames or vectors.  The package aims to provide
# functionality similar to that of SAS速 formats. Formats are assigned to
# the format attribute on data frame columns.  Then when the fdata()
# function is called, a new data frame is created with the column data
# formatted as specified.  The package also contains a value() function
# to create a user-defined format, similar to a SAS速 user-defined format.
# - Depends: R (>= 4.0.0), common
# - Linking To: Rcpp
# - Imports: tibble, stats, crayon, Rcpp
# - Suggests: testthat, knitr, rmarkdown, covr, logr, utils, dplyr, libr
# - Repository: CRAN
# - Functions: 49

```

```{r eval=FALSE, echo=TRUE}
# Get package info
res <- pkg_info("fmtr")

# View functions
print(res, verbose = TRUE)
# # A package info object: fmtr package
# - Version: v1.6.5
# - Release Date: 2024-06-13
# - Title: Easily Apply Formats to Data
# - Maintainer: David Bosak <dbosak01@gmail.com>
# - License: CC0
# - Description: Contains a set of functions that can be used to apply
# formats to data frames or vectors.  The package aims to provide
# functionality similar to that of SAS速 formats. Formats are assigned to
# the format attribute on data frame columns.  Then when the fdata()
# function is called, a new data frame is created with the column data
# formatted as specified.  The package also contains a value() function
# to create a user-defined format, similar to a SAS速 user-defined format.
# - Depends: R (>= 4.0.0), common
# - Linking To: Rcpp
# - Imports: tibble, stats, crayon, Rcpp
# - Suggests: testthat, knitr, rmarkdown, covr, logr, utils, dplyr, libr
# - Repository: CRAN
# - Functions: 
#   - as.data.frame.fcat(): x, row.names, optional, ...
#   - as.data.frame.fmt(): x, row.names, optional, ..., name
#   - as.data.frame.fmt_lst(): x, row.names, optional, ...
#   - as.fcat(): x
#   - as.fcat.data.frame(): x
#   - as.fcat.fmt_lst(): x
#   - as.fcat.list(): x
#   - as.fcat.tbl_df(): x
#   - as.flist(): x, type, lookup, simplify
#   - as.flist.data.frame(): x, type, lookup, simplify
#   - as.flist.fcat(): x, type, lookup, simplify
#   - as.flist.list(): x, type, lookup, simplify
#   - as.flist.tbl_df(): x, type, lookup, simplify
#   - as.fmt(): x
#   - as.fmt.data.frame(): x
#   - condition(): expr, label, order
#   - descriptions(): x
#   - descriptions<-(): x, value
#   - fapply(): x, format, width, justify
#   - fapply2(): x1, x2, format1, format2, sep, width, justify
#   - fattr(): x, format, width, justify, label, description, keep
#   - fattr<-()
#   - fcat(): ..., log
#   - fdata(): x, ...
#   - flist(): ..., type, lookup, simplify
#   - fmt_cnt_pct(): x, denom, format, na, zero
#   - fmt_mean_sd(): x, format, sd_format
#   - fmt_median(): x, format
#   - fmt_n(): x
#   - fmt_quantile_range(): x, format, sep, lower, upper, type
#   - fmt_range(): x, format, sep
#   - formats(): x
#   - formats<-(): x, value
#   - is.fcat(): x
#   - is.flist(): x
#   - is.format(): x
#   - justification(): x
#   - justification<-(): x, value
#   - labels.fmt(): object, ...
#   - print.fcat(): x, ..., verbose
#   - print.fmt(): x, ..., name, verbose
#   - print.fmt_lst(): x, ..., verbose
#   - read.fcat(): file_path
#   - read.flist(): file_path
#   - value(): ..., log, as.factor
#   - widths(): x
#   - widths<-(): x, value
#   - write.fcat(): x, dir_path, file_name
#   - write.flist(): x, dir_path, file_name

```


## Stability Assessment


```{r eval=FALSE, echo=TRUE}

# View package stability
pkg_stability("fmtr")
# # A stability score: fmtr package
# - Age: 4.50 years
# - Score: 96.5
# - Assessment: Very Stable
# - Version Range: 1.0.1/1.6.8
# - Release Range: 2020-08-07/2025-01-28
# - Release Count: 20
# - Breaking Releases: 1
# - Data: 
#    Package Version          FileName    Release   Size AF AP RF RP BC TF
# 1     fmtr   1.6.8 fmtr_1.6.8.tar.gz 2025-01-28 576.5K  0  1  0  0  0 49
# 2     fmtr   1.6.5 fmtr_1.6.5.tar.gz 2024-06-14   575K  0  0  0  0  0 49
# 3     fmtr   1.6.4 fmtr_1.6.4.tar.gz 2024-05-10   580K  2  0  0  0  0 49
# 4     fmtr   1.6.3 fmtr_1.6.3.tar.gz 2024-01-09   574K  0  0  0  0  0 47
# 5     fmtr   1.6.2 fmtr_1.6.2.tar.gz 2023-11-10   487K  0  1  0  0  0 47
# 6     fmtr   1.6.0 fmtr_1.6.0.tar.gz 2023-09-10   486K  0  0  0  0  0 47
# 7     fmtr   1.5.9 fmtr_1.5.9.tar.gz 2022-09-08   487K  0  3  0  0  0 47
# 8     fmtr   1.5.8 fmtr_1.5.8.tar.gz 2022-08-08   419K  1  1  2  0  1 47
# 9     fmtr   1.5.7 fmtr_1.5.7.tar.gz 2022-06-25   335K  0  0  0  0  0 48
# 10    fmtr   1.5.5 fmtr_1.5.5.tar.gz 2022-01-14   337K  0  1  0  0  0 48
# 11    fmtr   1.5.4 fmtr_1.5.4.tar.gz 2021-11-18   338K  0  0  0  0  0 48
# 12    fmtr   1.5.3 fmtr_1.5.3.tar.gz 2021-10-09   232K  0  0  0  0  0 48
# 13    fmtr   1.5.2 fmtr_1.5.2.tar.gz 2021-07-25   230K  0  0  0  0  0 48
# 14    fmtr   1.5.1 fmtr_1.5.1.tar.gz 2021-06-21   229K  0  0  0  0  0 48
# 15    fmtr   1.5.0 fmtr_1.5.0.tar.gz 2021-02-22   227K  2  0  0  0  0 48
# 16    fmtr   1.4.1 fmtr_1.4.1.tar.gz 2021-01-06   224K  0  0  0  0  0 46
# 17    fmtr   1.3.0 fmtr_1.3.0.tar.gz 2020-11-16    87K  6  1  0  0  0 46
# 18    fmtr   1.2.2 fmtr_1.2.2.tar.gz 2020-10-10    82K 17  0  0  0  0 40
# 19    fmtr   1.1.0 fmtr_1.1.0.tar.gz 2020-09-11    65K  6  1  0  0  0 23
# 20    fmtr   1.0.1 fmtr_1.0.1.tar.gz 2020-08-07    59K 17 17  0  0  0 17

```

## Comparing Package Versions


```{r eval=FALSE, echo=TRUE}
# Compare package versions
pkg_diff("fmtr", "1.5.7", "1.5.8")
# # A difference object: fmtr package
# - Comparing: v1.5.7/v1.5.8
# - Breaking Changes: TRUE
# - Added Functions: 
#   - fapply2()
# - Added Parameters: 
#   - fmt_cnt_pct(): na, zero
# - Removed Functions: 
#   - labels.data.frame()
#   - labels<-()

```



## Listing Package Versions


```{r eval=FALSE, echo=TRUE}

# Create vector of packages
v1 <- c("curl", "dplyr", "fmtr", "purrr")

# Get current versions
res1 <- pkg_repo(v1, ver = "current")
#   Package Version
# 1    curl   5.2.1
# 2   dplyr   1.1.4
# 3    fmtr   1.6.5
# 4   purrr   1.0.2

# Get latest versions
res2 <- pkg_repo(v1, ver = "latest")
#   Package Version
# 1    curl   6.2.0
# 2   dplyr   1.1.4
# 3    fmtr   1.6.8
# 4   purrr   1.0.4

```

## Anticipating Breakages



```{r eval=FALSE, echo=TRUE}
# View upgrade breakages
repo_breakages(res1, res2) 
# Comparing curl v5.2.1/v6.2.0
# Comparing fmtr v1.6.5/v1.6.8
# Comparing purrr v1.0.2/v1.0.4
# $summary
#   Package Version1 Version2 Breakages
# 1    curl    5.2.1    6.2.0      TRUE
# 2   dplyr    1.1.4    1.1.4     FALSE
# 3    fmtr    1.6.5    1.6.8     FALSE
# 4   purrr    1.0.2    1.0.4     FALSE
# 
# $details
# $details$curl
# # A difference object: curl package
# - Comparing: v5.2.1/v6.2.0
# - Breaking Changes: TRUE
# - Added Functions: 
#   - curl_parse_url()
# - Added Parameters: 
#   - multi_download(): multi_timeout
#   - multi_set(): max_streams
#   - new_pool(): max_streams
# - Removed Parameters: 
#   - multi_download(): timeout


```



## Assessing Entire Repository


```{r eval=FALSE, echo=TRUE}

repo_stability(v1)
#   Package    FV    LV         FR         LR TR BR     Score        Assessment
# 1    curl   0.2 6.2.0 2014-11-20 2025-01-23 50  1 0.9793617       Very Stable
# 2   dplyr   0.1 1.1.4 2014-01-16 2023-11-17 45 20 0.8711429 Somewhat Unstable
# 3    fmtr 1.0.1 1.6.8 2020-08-07 2025-01-28 20  1 0.9650000       Very Stable
# 4   purrr 0.1.0 1.0.4 2015-09-28 2025-02-05 18  6 0.9183333            Stable

```


