---
title: "Package Cache"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Package Cache}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pkgdiff)
```

To improve performance of difference operations, and especially 
of stability scores, **pkgdiff** uses a cache of pre-processed package
information on Github.  This cache greatly speeds up the operation
of **pkgdiff** functions.  The cache is created by a periodic process
that extracts information from packages on CRAN, and stores it in a way
that the information can be quickly retrieved when a **pkgdiff** function
is called.

Note the following about the **pkgdiff** cache:
1) To access the cache, the computer **pkgdiff** is running on must have 
access to the internet and, specifically, Github.  If your machines is 
isolated from the internet, or blocked from Github, **pkgdiff** will not run.
2) The cache does not contain all CRAN packages.  It contains packages that
exceed 1000 downloads per month.  These packages constitute approximately 16%
of the top CRAN downloads, or about 3500 packages. For most R users, any package 
you might be interested in will be included in this subset.  
3) If a package is not in the cache, all **pkgdiff** functions will still 
work on that package.  However, they will be much slower than packages
that are included in the cache. 
4) You may request that a package be added to the cache by submitting 
an issue to the **pkgdiff** issue log 
[here](https://github.com/dbosak01/pkgdiff/issues).
5) Also observe that the `pkg_info()` function returns a parameter
on whether or not the selected package is in the cache.  It is recommended
to run `pkg_info()` first, and check if the package is in the cache before
running `pkg_stability()`.
6) To find out which packages are in the **pkgdiff** cache, 
you may call the function `pkg_cache()`. See the documentation of that
function for additional information.

