---
title: "pkgdiff"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pkgdiff}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pkgdiff)
```
The purpose of **pkgdiff** is to identify changes between package versions,
with the aim of assessing package stability.  Here is a list of the major
functions in **pkgdiff**, and a short explanation of each:

- `pkg_info()`: Displays general information about a package,
including the description, maintainer, and number of downloads in
the last month.
- `pkg_stability()`: Looks at the history of a package,
identifies breaking changes, and assesses the package for stability.
- `pkg_diff()`: Compares two versions of a package, and shows
which functions and parameters have been added or removed.
- `pkg_repo()`: Generates a list of the packages in a repository
and their versions.
- `repo_breakages()`: Looks at a package repository and tries to identify
which packages will break when upgraded.
- `repo_stability()`: Looks at a package repository and calculates
a stability score for each package.

Below is a further explanation and demo for each of the above functions.

## General Package Information

The first thing you may want to do when exploring a package is to get
some general information about it.  General information about a package
can be achieved using the `pkg_info()` function.  All you need to 
pass is the package name. The function will then return basic 
information such as the maintainer, description,
and imported packages.  Here is a simple demonstration:

```{r eval=FALSE, echo=TRUE}
library(pkgdiff)

# Examine tidymodels package
pkg_info("tidymodels")
# # A package info object: tidymodels package
# - Version: v1.2.0
# - Release Date: 2024-03-25
# - Title: Easily Install and Load the 'Tidymodels' Packages
# - Maintainer: Max Kuhn <max@posit.co>
# - License: MIT + file LICENSE
# - Description: The tidy modeling "verse" is a collection of packages for
# modeling and statistical analysis that share the underlying design
# philosophy, grammar, and data structures of the tidyverse.
# - Depends: R (>= 3.6)
# - Imports: broom (>= 1.0.5), cli (>= 3.6.2), conflicted (>= 1.2.0), dials (>= 1.2.1), dplyr (>= 1.1.4), ggplot2 (>= 3.5.0), hardhat (>= 1.3.1), infer (>= 1.0.6), modeldata (>= 1.3.0), parsnip (>= 1.2.1), purrr (>= 1.0.2), recipes (>= 1.0.10), rlang (>= 1.1.3), rsample (>= 1.2.1), rstudioapi (>= 0.16.0), tibble (>= 3.2.1), tidyr (>= 1.3.1), tune (>= 1.2.0), workflows (>= 1.1.4), workflowsets (>= 1.1.0), yardstick (>= 1.3.1)
# - Suggests: covr, glue, knitr, rmarkdown, testthat (>= 3.0.0), xml2
# - Repository: CRAN
# - Cached: TRUE
# - Functions: 9

```
The return value of this function is actually an object that can be manipulated
programmatically.  Here we will extract some of the data values from this object:

```{r eval=FALSE, echo=TRUE}
# Capture package info object
res <- pkg_info("tidymodels")

# View version
res$Version
# [1] "1.3.0"

# View title
res$Title
# [1] "Easily Install and Load the 'Tidymodels' Packages"

# View functions
res$ReleaseDate
# [1] "2025-02-21"

```
Note that, by default, the version of the package returned is the version of the package
located on the current machine, in the current version of the R repository.
If the package is not installed on in the current R respository, the function
will return the latest version on CRAN.

Also note that the object contains a complete list of all exported package 
functions. This list can be accessed programatically, and also viewed using
the `verbose` parameter of the `print()` function:

```{r eval=FALSE, echo=TRUE}
# Get package info
res <- pkg_info("tidymodels")

# View functions
print(res, verbose = TRUE)
# # A package info object: tidymodels package
# - Version: v1.3.0
# - Release Date: 2025-02-21
# - Title: Easily Install and Load the 'Tidymodels' Packages
# - Maintainer: Max Kuhn <max@posit.co>
# - License: MIT + file LICENSE
# - Description: The tidy modeling "verse" is a collection of packages for
# modeling and statistical analysis that share the underlying design
# philosophy, grammar, and data structures of the tidyverse.
# - Depends: R (>= 3.6)
# - Imports: broom (>= 1.0.7), cli (>= 3.6.4), conflicted (>= 1.2.0), dials (>= 1.4.0), dplyr (>= 1.1.4), ggplot2 (>= 3.5.1), hardhat (>= 1.4.1), infer (>= 1.0.7), modeldata (>= 1.4.0), parsnip (>= 1.3.0), purrr (>= 1.0.4), recipes (>= 1.1.1), rlang (>= 1.1.5), rsample (>= 1.2.1), rstudioapi (>= 0.17.1), tibble (>= 3.2.1), tidyr (>= 1.3.1), tune (>= 1.3.0), workflows (>= 1.2.0), workflowsets (>= 1.1.0), yardstick (>= 1.3.2)
# - Suggests: covr, glue, knitr, rmarkdown, testthat (>= 3.0.0), xml2
# - Downloads/Month: 31789
# - Repository: CRAN
# - Cached: TRUE
# - Functions: 
#   - pkg_deps(): x, recursive
#   - print.tidymodels_conflicts(): x, ..., startup
#   - tag_attach(): tag
#   - tag_show()
#   - tag_update(): tag
#   - tidymodels_conflicts()
#   - tidymodels_packages(): include_self
#   - tidymodels_prefer(): quiet
#   - tidymodels_update(): pkg, recursive, ...
```

## Stability Assessment

Besides general information about a package, you may also view the "stability"
of a package.  The `pkg_stability()` function looks at the entire history
of the package, and attempts to assess its overall stability in terms of 
breaking changes.  Here, a "breaking change" means either a removed function
or removed function parameter.  This function is very useful when you are
writing production code, and want to select the most stable packages possible.
The results of this function can help you make that determination: 

```{r eval=FALSE, echo=TRUE}

# View package stability
pkg_stability("tidymodels")
# # A stability score: tidymodels package
# - Age: 6.58 years
# - Score: 100.0
# - Assessment: Perfect
# - Version Range: 0.0.1/1.3.0
# - Release Range: 2018-07-27/2025-02-21
# - Release Count: 14
# - Breaking Releases: 0
# - Data: 
#       Package Version                FileName    Release  Size AF AP RF RP BC TF
# 1  tidymodels   1.3.0 tidymodels_1.3.0.tar.gz 2025-02-21 46.4K  0  0  0  0  0  9
# 2  tidymodels   1.2.0 tidymodels_1.2.0.tar.gz 2024-03-25 49.6K  0  0  0  0  0  9
# 3  tidymodels   1.1.1 tidymodels_1.1.1.tar.gz 2023-08-24   50K  0  0  0  0  0  9
# 4  tidymodels   1.1.0 tidymodels_1.1.0.tar.gz 2023-05-01   49K  0  0  0  0  0  9
# 5  tidymodels   1.0.0 tidymodels_1.0.0.tar.gz 2022-07-13   75K  0  0  0  0  0  9
# 6  tidymodels   0.2.0 tidymodels_0.2.0.tar.gz 2022-03-19   75K  0  0  0  0  0  9
# 7  tidymodels   0.1.4 tidymodels_0.1.4.tar.gz 2021-10-01   17K  0  0  0  0  0  9
# 8  tidymodels   0.1.3 tidymodels_0.1.3.tar.gz 2021-04-19   15K  1  0  0  0  0  9
# 9  tidymodels   0.1.2 tidymodels_0.1.2.tar.gz 2020-11-22   26K  0  0  0  0  0  8
# 10 tidymodels   0.1.1 tidymodels_0.1.1.tar.gz 2020-07-14   25K  0  0  0  0  0  8
# 11 tidymodels   0.1.0 tidymodels_0.1.0.tar.gz 2020-02-16   57K  0  0  0  0  0  8
# 12 tidymodels   0.0.3 tidymodels_0.0.3.tar.gz 2019-10-05   19K  0  0  0  0  0  8
# 13 tidymodels   0.0.2 tidymodels_0.0.2.tar.gz 2018-11-27   18K  0  0  0  0  0  8
# 14 tidymodels   0.0.1 tidymodels_0.0.1.tar.gz 2018-07-27   18K  8  8  0  0  0  8

```
The stability score is calculated generally as the percentage of non-breaking
changes.  Breaking changes are also weighted to reduce their impact over time.
The above score of 100 for **tidymodels** means that the package has never
had a breaking change.  It is therefore given a stability assessment of "Perfect".

Now let's look at a package with more breaking changes:
```{r eval=FALSE, echo=TRUE}

# View package stability
pkg_stability("dbplyr")
# # A stability score: dbplyr package
# - Age: 7.67 years
# - Score: 73.7
# - Assessment: Unstable
# - Version Range: 1.0.0/2.5.0
# - Release Range: 2017-06-09/2024-03-19
# - Release Count: 23
# - Breaking Releases: 10
# - Data: 
#    Package Version            FileName    Release   Size  AF  AP RF RP BC  TF
# 1   dbplyr   2.5.0 dbplyr_2.5.0.tar.gz 2024-03-19 752.6K  32   6 12  0  1 611
# 2   dbplyr   2.4.0 dbplyr_2.4.0.tar.gz 2023-10-26   743K  74  80 45 33  1 591
# 3   dbplyr   2.3.4 dbplyr_2.3.4.tar.gz 2023-09-27   706K   0   0  0  0  0 562
# 4   dbplyr   2.3.3 dbplyr_2.3.3.tar.gz 2023-07-07   706K   0   0  0  0  0 562
# 5   dbplyr   2.3.2 dbplyr_2.3.2.tar.gz 2023-03-21   707K   0   1  0  1  1 562
# 6   dbplyr   2.3.1 dbplyr_2.3.1.tar.gz 2023-02-24   707K   5   6  1  0  1 562
# 7   dbplyr   2.3.0 dbplyr_2.3.0.tar.gz 2023-01-16   688K  33  16  5  1  1 558
# 8   dbplyr   2.2.1 dbplyr_2.2.1.tar.gz 2022-06-27   655K   4   0  0  0  0 530
# 9   dbplyr   2.2.0 dbplyr_2.2.0.tar.gz 2022-06-05   655K 136  47 53  1  1 526
# 10  dbplyr   2.1.1 dbplyr_2.1.1.tar.gz 2021-04-06   568K   5   0  0  0  0 443
# 11  dbplyr   2.1.0 dbplyr_2.1.0.tar.gz 2021-02-03   566K  15   2  0  0  0 438
# 12  dbplyr   2.0.0 dbplyr_2.0.0.tar.gz 2020-11-03   522K 139  33 94 11  1 423
# 13  dbplyr   1.4.4 dbplyr_1.4.4.tar.gz 2020-05-27   417K   5   0  0  0  0 378
# 14  dbplyr   1.4.3 dbplyr_1.4.3.tar.gz 2020-04-19   413K   9   3  2  2  1 373
# 15  dbplyr   1.4.2 dbplyr_1.4.2.tar.gz 2019-06-17   296K   1   0  0  0  0 366
# 16  dbplyr   1.4.1 dbplyr_1.4.1.tar.gz 2019-06-06   296K   1   0  0  0  0 365
# 17  dbplyr   1.4.0 dbplyr_1.4.0.tar.gz 2019-04-24   295K  24  30 17 13  1 364
# 18  dbplyr   1.3.0 dbplyr_1.3.0.tar.gz 2019-01-09   263K   2   0  0  0  0 357
# 19  dbplyr   1.2.2 dbplyr_1.2.2.tar.gz 2018-07-25   258K   0   0  0  0  0 355
# 20  dbplyr   1.2.1 dbplyr_1.2.1.tar.gz 2018-02-19   257K   0   0  0  0  0 355
# 21  dbplyr   1.2.0 dbplyr_1.2.0.tar.gz 2018-01-03   257K  43   6  2  2  1 355
# 22  dbplyr   1.1.0 dbplyr_1.1.0.tar.gz 2017-06-27   241K  23   2  0  0  0 314
# 23  dbplyr   1.0.0 dbplyr_1.0.0.tar.gz 2017-06-09   235K 291 291  0  0  0 291

```
As seen above, the **dbplyr** package has had a lot of breaking changes.  Which releases
had breaking changes are shown the the "BC" column of the data grid.  The "AF" and
"AP" columns stand for "Added Functions" and "Added Parameters", while
"RF" and "RP" mean "Removed Functions" and "Removed Parameters".  "TF" is the total
number of exported functions in that version of the package.

Since this package has issued breaking changes in 10 out of 23 releases, it gets
a stability score of 73.7 and an assessment of "Unstable".  A stability assessment
of "Unstable" does not necessarily mean you should not use the package.  Whether
or not you use the package depends on your situation.  The `pkg_stability()` 
function merely contributes information on which you can make your own decisions.

## Comparing Package Versions

If you find a package with breaking changes, you would then want to know
what those breaking changes are.  The breaking changes between package versions
can be discovered with the `pkg_diff()` function.  This function accepts 
a package name, and the two version of the package you wish to compare.  
Looking at our stability data above, let's examine the breaking changes 
between versions "2.4.0" and "2.5.0" of **dbplyr**:

```{r eval=FALSE, echo=TRUE}
# Compare package versions
pkg_diff("dbplyr", "2.4.0", "2.5.0")
# # A difference object: dbplyr package
# - Comparing: v2.4.0/v2.5.0
# - Breaking Changes: TRUE
# - Added Functions: 
#   - [.dbplyr_table_path()
#   - [[.dbplyr_table_path()
#   - as_table_path()
#   - c.dbplyr_table_path()
#   - check_table_path()
#   - complete.tbl_lazy()
#   - db_explain.Oracle()
#   - db_explain.OraConnection()
#   - db_sql_render.Microsoft SQL Server()
#   - escape.dbplyr_table_path()
#   - escape.default()
#   - expand.tbl_lazy()
#   - fill.tbl_lazy()
#   - group_by_drop_default.tbl_lazy()
#   - intersect.tbl_lazy()
#   - is_table_path()
#   - pivot_longer.tbl_lazy()
#   - pivot_wider.tbl_lazy()
#   - print.dbplyr_table_path()
#   - reframe.tbl_lazy()
#   - replace_na.tbl_lazy()
#   - setdiff.OraConnection()
#   - setdiff.tbl_lazy()
#   - setdiff.tbl_Oracle()
#   - simulate_mariadb()
#   - sql_escape_ident.default()
#   - sql_escape_ident.MySQLConnection()
#   - sql_escape_string.default()
#   - table_path_components()
#   - table_path_components.default()
#   - table_path_name()
#   - union.tbl_lazy()
# - Added Parameters: 
#   - flatten_query.base_query(): con
#   - flatten_query.multi_join_query(): con
#   - flatten_query.select_query(): con
#   - flatten_query.union_query(): con
#   - flatten_query.values_query(): con
#   - sql_translation.MySQL(): con
# - Removed Functions: 
#   - as_table_ident.character()
#   - as_table_ident.dbplyr_catalog()
#   - as_table_ident.dbplyr_schema()
#   - as_table_ident.dbplyr_table_ident()
#   - as_table_ident.Id()
#   - as_table_ident.ident()
#   - as_table_ident.ident_q()
#   - as_table_ident.sql()
#   - escape.dbplyr_table_ident()
#   - format.dbplyr_table_ident()
#   - sql_escape_ident.DBIConnection()
#   - sql_escape_string.DBIConnection()

```
As you can see, many new functions and function parameter have been added
between these two versions.  In addition, there are 12 functions that 
have been removed.  It is up to you to determine whether and how these 
removals will affect your programs.  It is very convenient, however,
to have an accurate accounting of package changes.  These changes
can help you plan your version upgrades and anticipate breakages.


## Listing Package Versions


```{r eval=FALSE, echo=TRUE}

# Create vector of packages
v1 <- c("curl", "dplyr", "fmtr", "purrr")

# Get current versions
res1 <- pkg_repo(v1, ver = "current")
#   Package Version
# 1    curl   5.2.1
# 2   dplyr   1.1.4
# 3    fmtr   1.6.5
# 4   purrr   1.0.2

# Get latest versions
res2 <- pkg_repo(v1, ver = "latest")
#   Package Version
# 1    curl   6.2.0
# 2   dplyr   1.1.4
# 3    fmtr   1.6.8
# 4   purrr   1.0.4

```

## Anticipating Breakages


### Small list of packages

```{r eval=FALSE, echo=TRUE}
# View upgrade breakages
repo_breakages(res1, res2) 
# Comparing curl v5.2.1/v6.2.0
# Comparing fmtr v1.6.5/v1.6.8
# Comparing purrr v1.0.2/v1.0.4
# $summary
#   Package Version1 Version2 Breakages
# 1    curl    5.2.1    6.2.0      TRUE
# 2   dplyr    1.1.4    1.1.4     FALSE
# 3    fmtr    1.6.5    1.6.8     FALSE
# 4   purrr    1.0.2    1.0.4     FALSE
# 
# $details
# $details$curl
# # A difference object: curl package
# - Comparing: v5.2.1/v6.2.0
# - Breaking Changes: TRUE
# - Added Functions: 
#   - curl_parse_url()
# - Added Parameters: 
#   - multi_download(): multi_timeout
#   - multi_set(): max_streams
#   - new_pool(): max_streams
# - Removed Parameters: 
#   - multi_download(): timeout
```

### Entire Repository
```{r eval=FALSE, echo=TRUE}

# Get current versions from local repo
r1 <- pkg_repo(ver = "current")

# View Results
r1
#            Package              Version
# 1            abind                1.4-5
# 2          admiral               0.10.1
# 3       admiraldev                0.3.0
# 4      admiralonco                0.3.0
# 5             ards                0.1.0
# 6            arrow             11.0.0.3
# 7          askpass                  1.1
# 8       assertthat                0.2.1
# 9           aws.s3               0.3.21
# 10   aws.signature                0.6.0
# 11       backports                1.4.1
# 12       base64enc                0.1-3
# 13             bit                4.0.4
# 14           bit64                4.0.5
# 15          bitops                1.0-7
# 16            blob                1.2.3
# ...

# Get latest versions of current repo
r2 <- pkg_repo(r1$Package, ver = "latest")

# View Results
r2
#            Package    Version
# 1            abind      1.4-8
# 2          admiral      1.2.0
# 3       admiraldev      1.2.0
# 4      admiralonco      1.2.0
# 5             ards      0.1.1
# 6            arrow   18.1.0.1
# 7          askpass      1.2.1
# 8       assertthat      0.2.1
# 9           aws.s3     0.3.21
# 10   aws.signature      0.6.0
# 11       backports      1.5.0
# 12       base64enc      0.1-3
# 13             bit    4.5.0.1
# 14           bit64    4.6.0-1
# 15          bitops      1.0-9
# 16            blob      1.2.4
# ...


# Check for breakages on whole repository
res <- repo_breakages(r1, r2)
# Comparing abind v1.4-5/v1.4-8
# Comparing admiral v0.10.1/v1.2.0
# Comparing admiraldev v0.3.0/v1.2.0
# Comparing admiralonco v0.3.0/v1.2.0
# Comparing ards v0.1.0/v0.1.1
# Archive versions of 'ards' not available.
# Comparing arrow v11.0.0.3/v18.1.0.1
# Comparing askpass v1.1/v1.2.1
# Comparing backports v1.4.1/v1.5.0
# Comparing bit v4.0.4/v4.5.0.1
# Comparing bit64 v4.0.5/v4.6.0-1
# Comparing bitops v1.0-7/v1.0-9
# Comparing blob v1.2.3/v1.2.4
# Comparing box v1.1.3/v1.2.0
# Comparing brew v1.0-8/v1.0-10
# Comparing brio v1.1.3/v1.1.5
# Comparing broom v1.0.1/v1.0.7
# Comparing bslib v0.4.0/v0.9.0
# Comparing cachem v1.0.6/v1.1.0
# Comparing callr v3.7.3/v3.7.6
```



```{r eval=FALSE, echo=TRUE}
# View Summary Table
res$summary
#            Package             Version1   Version2 Breakages
# 1            abind                1.4-5      1.4-8     FALSE
# 2          admiral               0.10.1      1.2.0      TRUE
# 3       admiraldev                0.3.0      1.2.0      TRUE
# 4      admiralonco                0.3.0      1.2.0      TRUE
# 5             ards                0.1.0      0.1.1        NA
# 6            arrow             11.0.0.3   18.1.0.1     FALSE
# 7          askpass                  1.1      1.2.1     FALSE
# 8       assertthat                0.2.1      0.2.1     FALSE
# 9           aws.s3               0.3.21     0.3.21     FALSE
# 10   aws.signature                0.6.0      0.6.0     FALSE
# 11       backports                1.4.1      1.5.0     FALSE
# 12       base64enc                0.1-3      0.1-3     FALSE
# 13             bit                4.0.4    4.5.0.1     FALSE
# 14           bit64                4.0.5    4.6.0-1      TRUE
# 15          bitops                1.0-7      1.0-9     FALSE
# 16            blob                1.2.3      1.2.4     FALSE
# ...

# View Difference Details
res$details$admiralonco
# # A difference object: admiralonco package
# - Comparing: v0.3.0/v1.2.0
# - Breaking Changes: TRUE
# - Added Functions: 
#   - bor_cr()
#   - bor_ne()
#   - bor_non_crpd()
#   - bor_pd()
#   - bor_pr()
#   - bor_sd()
#   - cb_y()
#   - cbor_cr()
#   - cbor_pr()
#   - crsp_y_cr()
#   - crsp_y_pr()
#   - no_data_missing()
#   - no_data_n()
#   - rsp_y()
# - Removed Functions: 
#   - call_aval_fun()
#   - derive_param_lasta()

```



## Assessing Entire Repository


```{r eval=FALSE, echo=TRUE}

repo_stability(v1)
#   Package    FV    LV         FR         LR TR BR     Score        Assessment
# 1    curl   0.2 6.2.0 2014-11-20 2025-01-23 50  1 0.9793617       Very Stable
# 2   dplyr   0.1 1.1.4 2014-01-16 2023-11-17 45 20 0.8711429 Somewhat Unstable
# 3    fmtr 1.0.1 1.6.8 2020-08-07 2025-01-28 20  1 0.9650000       Very Stable
# 4   purrr 0.1.0 1.0.4 2015-09-28 2025-02-05 18  6 0.9183333            Stable

```


